---
title: |
    | S2Y Lab 4
    | How good is the model? – $R^2$ and residual plots
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
---

```{r setup, include = FALSE}
library(webexercises)
library(tinytex)
library(dplyr)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)

diamonds <- read.csv("diamonds.csv")
```


```{r include=FALSE, echo=FALSE}
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```


# Welcome to S2Y Lab 4

Intended Learning Outcomes:

- Assess the goodness of fit of linear models based on $R^2$ and $R^2$ (adj);
- Produce residual plots; 
- Assess the assumptions of a linear model using residual plots.

## Introduction {-#intro}

In the lectures $R^2$ was used to measure the goodness-of-fit of the chosen model, and the idea of residuals was introduced. In today's practical we will use RStudio to find values of $R^2$, and use residual plots to examine the assumptions underlying the linear model.

Similarly to previous practicals, we will use the `lm()` command to fit linear models. We will obtain residuals and standardised residuals from linear models that allow us to check assumptions. We will check our assumptions using:

(i) plots of residuals versus fitted values;
(ii) Normal Q-Q plots;
(iii) histograms of the residuals.
